---
title: "Evanston Arrests"
author: "Patrick Tu"
date: "11/6/2022"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  error = FALSE,
  message = FALSE,
  warning = FALSE)
```

For our next R analysis we decided to focus once again on the city of Evanston, Illinois. This time we are focusing on the demographics of arrests in Evanston using [this](https://data.cityofevanston.org/Police/Evanston-Arrests/25em-v4fn) data set from the City of Evanston database. We wanted to know what races, ages, and genders were the most likely to be arrested in Evanston. 

```{r}
library("RSocrata")
library("tidyverse")

arrests <- 
  as_tibble(
    read.socrata("https://data.cityofevanston.org/resource/25em-v4fn.json")
    ) %>%  
  mutate(
      day_of_the_week = factor(day_of_the_week),
      day_of_the_week = fct_relevel(
        day_of_the_week, 
        values = c(
          "Mon", 
          "Tue", 
          "Wed", 
          "Thu", 
          "Fri", 
          "Sat", 
          "Sun" 
        )
      )
    ) %>%
  separate(
    col = arrest_date, 
    into = c("Year", "Month", "Date"), 
    remove = FALSE
    ) %>%
  mutate(Year = as.numeric(Year), 
         age = as.numeric(age), 
         arrest_time = as.numeric(arrest_time),
         street_number = as.numeric(street_number)) %>%
  filter(Year > 2016)
```  

We were interested in how arrests have changed over the years and if arrests happen more on certain days of the week so we graphed those variables here

```{r}
ggplot(arrests, aes(x = day_of_the_week))+
  geom_bar()+
  facet_wrap(~Year)

ggplot(arrests, aes(x = Year))+
  geom_bar()

arrests %>%
  count(Year)

ggplot(arrests, aes(x = Month))+
  geom_bar() +
  facet_wrap(~Year, nrow = 1)
```



```{r}
chi_data_df <- 
  arrests %>% 
  count(Year, Month) %>%
  pivot_wider(
    id_cols = Month, 
    names_from = Year, 
    values_from = n, 
    names_prefix = "m_"
    )

chi_data_mat <- chi_data_df %>% select(-Month) %>% as.matrix(.)
rownames(chi_data_mat) <- chi_data_df$Month

# chi-square test of independence
chisq.test(chi_data_mat) # significant!
```

Next we decided to plot graphs comparing age, race, and gender of those arrested

```{r}
ggplot(arrests%>%
         filter(complete.cases(sex)), aes(x = age))+
  geom_histogram(binwidth = 2)+
  coord_cartesian(xlim = c(0,90))+
  scale_x_continuous(breaks = seq(0,90,10))+
  facet_wrap(~sex)

arrests_gender <- arrests %>%
  filter(complete.cases(sex, age)) %>%
  group_by(sex) %>%
  summarise(
    gender_mean = mean(age), 
    gender_median = median(age), 
    gender_sd = sd(age), 
    gender_sample = n(), 
    gender_min = min(age), 
    gender_max = max(age)
    )

t.test(age~sex, data = arrests, alternative = "two.sided")

arrests_race <- arrests %>%
  filter(complete.cases(race, age))%>%
  group_by(race) %>%
  summarise(
    race_mean = mean(age), 
    race_median = median(age), 
    race_sd = sd(age), 
    race_sample = n(), 
    race_min = min(age), 
    race_max = max(age)
  )

ggplot(arrests_race, aes(x = race, y = race_mean))+
  geom_point(
    data = arrests%>%filter(complete.cases(race, age)), 
    aes(y = age), 
    position = position_jitter(.22), 
    alpha = 1/2, 
    color = "lightgrey", 
    shape = 1)+
  geom_errorbar(
    aes(ymin = race_mean - race_sd, ymax = race_mean + race_sd), 
    width = 0.2, color = "darkolivegreen")+
  geom_point(aes(size = race_sample), color = "darkolivegreen")+
  theme_bw()+
  labs(
    x = "Race", 
    y = "Average Age", 
    caption = "Error bars are sd", 
    title = "Arrests in Evanston by Age and Race"
    )

arrests_race_gender <- arrests %>%
  group_by(sex, race) %>%
  summarise(racesex_mean = mean(age, na.rm = TRUE))

ggplot(arrests_gender, aes(x = sex, y = gender_mean))+
  geom_point(
    data = arrests%>%filter(complete.cases(race, age)), 
    aes(y = age), 
    position = position_jitter(.22), 
    alpha = 1/2, 
    color = "lightgrey", 
    shape = 1)+
  geom_errorbar(
    aes(ymin = gender_mean - gender_sd, ymax = gender_mean + gender_sd), 
    width = 0.2, color = "darkolivegreen")+
  geom_point(aes(size = gender_sample), color = "darkolivegreen")+
  theme_bw()+
  labs(
    x = "Sex", 
    y = "Average Age", 
    caption = "Error bars are sd", 
    title = "Arrests in Evanston by Age and Sex"
  )

arrests_final <- arrests%>%
  filter(complete.cases(race, sex))

ggplot(arrests_final, aes(x = age))+
  geom_histogram(binwidth = 2)+
  coord_cartesian(xlim = c(0,90))+
  scale_x_continuous(breaks = seq(0,90,10))+
 facet_grid(race~sex)
  
arrests_racesex <- arrests_final %>%
  filter(complete.cases(age))%>%
  group_by(race, sex) %>%
  summarise(
    m = mean(age), 
    med = median(age), 
    sd = sd(age), 
    n = n(), 
    min = min(age), 
    max = max(age)
  )%>%
  ungroup()

ggplot(arrests_racesex, aes(x = race, y = med))+
  geom_point(
    data = arrests_final, 
    aes(y = age), 
    position = position_jitter(.22), 
    alpha = 1/2, 
    color = "lightgrey", 
    shape = 1)+
  geom_errorbar(
    aes(ymin = med - sd, ymax = med + sd), 
    width = 0.2, color = "darkolivegreen")+
  geom_point(aes(size = n), color = "darkolivegreen")+
  theme_bw()+
  labs(
    x = "Race", 
    y = "Median Age", 
    caption = "Error bars are sd", 
    title = "Arrests in Evanston by Race, Sex, and Age"
  )+
  facet_wrap(~sex)
```

```{r}
arrests_bwmale <- arrests_final%>%
  filter(complete.cases(age))%>%
  filter(sex == "Male", race %in% c("White", "Black"))

t.test(age~race, data = arrests_bwmale, alternative = "two.sided")

ggplot(arrests, aes(x = arrest_time))+
  geom_histogram(binwidth = 5)

ggplot(arrests, aes(x = street_number))+
  geom_histogram(binwidth = 100)

arrests %>%
  count(street_number)%>%
  arrange(-n)
```
