---
title: "NBA Player Comparison Function"
author: "Patrick Tu"
date: "5/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  error  = FALSE,
  message = FALSE)
```

I developed a function that compares two players in the NBA using the `nbastatR` package. The function compares the players based on their advanced and/or per game statistics.

If you want to use this function it is on my github here is how you would source it onto your workspace.

*Notes:* If you want to use the function you will have to download the `nbastatR`, `dplyr`, `ggplot2`, `tidyr`, `Hmisc`, `ghibli` packages.

```{r}
library(devtools)
source_url("https://raw.githubusercontent.com/pltu06/pltu06.github.io/main/NBA_player_function.R")
```

Here's some different ways to use the function.

First, let's compare Jimmy Butler and Jaysom Tatum:

```{r}
player_comparison(
  player1 = "Jimmy Butler",
  player2 = "Jayson Tatum", 
  table = c("per_game", "advanced"),
  season = 2022
)
```

<br >
<br >

You can even change the year for comparison as long as both players played in that year. Let's look at 2021:

```{r}
player_comparison(
  player1 = "Jimmy Butler",
  player2 = "Jayson Tatum", 
  table = c("per_game", "advanced"),
  season = 2021
)
```

<br>
<br>

If you only want the per game statistics do this

```{r}
player_comparison(
  player1 = "Jimmy Butler",
  player2 = "Jayson Tatum", 
  table = c("per_game"),
  season = 2021
)
```
<br>
<br>

If you want to have only the advanced statistics do this

```{r}
player_comparison(
  player1 = "Jimmy Butler",
  player2 = "Jayson Tatum", 
  table = c("advanced"),
  season = 2021
)
```

Here's the code that goes into that function. You can also see it on github [here](https://github.com/pltu06/pltu06.github.io/blob/main/NBA_player_function.R)

```{r, echo=FALSE}
library(knitr)
spin_child("NBA_player_function.R")
```
<br>